---
title: "field_dataset_analysis"
author: "Jianqiu Zheng"
date: "11/28/2020"
output:
  pdf_document: default
  html_document: default
---

## dataset prep
This section of code extract unique soil samples into a new dataframe <rate> 

```{r dataset}
library(ggplot2)
library(reshape2)
library(dplyr)
library(ggpubr)
library(GGally)
library(scales)
library(MASS)


# Create a transparent theme object
transparent_theme <- theme(
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  axis.text.x = element_blank(), 
  axis.text.y = element_blank(),
  axis.ticks = element_blank(),
  panel.grid = element_blank(),
  axis.line = element_blank(),
  panel.background = element_rect(fill = "transparent",colour = NA),
  plot.background = element_rect(fill = "transparent",colour = NA))



fdata <- read.csv(file = 'field_dataset.csv') # dataset use a dummy colomn to scale all ra measurements


#==data fitting
library(minpack.lm)
x<-fdata$rs
y<-fdata$ra
dat<-data.frame(x=x)
dat$y<-y


curve.nls = nlsLM(y ~ a*x^2+b*x,
                     start=list(a=-3,
                                b=3),
                     data = dat)
coef(curve.nls)
summary(curve.nls)


#plot observations along clay and soc gradient
plot1<-ggplot(fdata, aes(x = rs,y = ra, group = id)) +
  scale_x_continuous(name = expression(paste("Relative saturation")),limits = c(0,1)) +
  scale_y_continuous(name = 'Relative activity ',limits = c(0,1)) +
  geom_point(aes(color=clay),size=1.6, shape=1,stroke=1)+theme_bw()+
  scale_color_gradientn(colors=c("#b35806", "#fee0b6", "#542788"), values=c(0,1))+theme(text = element_text(size=16)) 
plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))#co
plot2

rs<-seq(0,1,0.05)
femp<-(1.98*rs-1.51*rs^2)/0.65#/0.644  #max=0.644  (3.07RS-2.36RS^2)
emp<-data.frame(rs=rs)
emp$res<-femp  #relative activity

bp<-ggplot(emp, aes(x = rs,y = femp)) +geom_line(color="black",size=1,shape=1,stroke=1)+transparent_theme
bp_grob<-ggplotGrob(bp)#create the external graphcal elements (called a grop in Grid terminology)
ymin<-0
xmin<-0
plot7<-plot2+annotation_custom(grob=bp_grob, xmin =xmin+0.1, xmax=xmin+1, ymin=ymin, ymax=ymin+0.75)
print(plot7)

pdf("field_obv_clay.pdf", width=5.5, height=4)
plot7
dev.off() 

```
#==subsetting unique sample ====
```{r unique}
rate<-dplyr::select(filter(fdata,ra==1), c(id, rs, mvol, mp, bd, porosity, clay, org))

#===calculating accessible soc and oxy at optimum moisture condition (when ra==1)
opt<-function(po,mvol,bd,org){
  soc<-org/12*bd*1e-3  #assum ss 1% as DOC  (mol/cm3) and 10% of these DOC are accessible
  oxy<-(po-mvol)*0.21*1.3e-6  #(mol/cm3)Henry's law constant (Caq/p=1.3e-3 mol/L/atm)   
  return (list(soc, oxy))}
po<-rate$porosity
mvol<-rate$mvol
org<-rate$org
bd<-rate$bd
ss_opt<-opt(po,mvol,bd,org)
rate$soc_ss<-ss_opt[[1]] ##approximately 100mol/cm3
rate$oxy_ss<-ss_opt[[2]]
rate$mp_fc<-rate$mp
summary(rate)
```

## Effective Km calculation
This section prepares parameters for effective Km calculations (based on Tang et al. 2019) 
```{r Kaeff}
Npsite=3000 # number of transporter per cell
k2_p=100  #transporter specific substrate uptake rate, unit:s-1
rc=1.e-6 #microbial cell radius unit:m
rp=1.e-9  #transporter radius unit:m
Na=6.e23  #Avogadro number
Ratm=50 #atmospheric resistance, 50 s/m

temp=25+273.15#define temperature

#=calculating gaseous and aqueous tortuosity (Original paper based on Morldrup, 2003)
tau<-function (mvol, po, clay){
  wpo<-mvol  #water filled porosity
  gpo<-po-mvol #gas filled porosity
  b<-2.91+0.195*clay #shape parameter 
  taug<- (po-mvol)*((po-mvol)/po)^3/b
  tauw<- mvol*(mvol/po)^(3/b-1)
  return (list(wpo=wpo, gpo=gpo,taug=taug,tauw=tauw))
}
mvol<-rate$mvol
po<-rate$porosity
clay<-rate$clay
tortuo<-tau(mvol, po,clay)

#==calculating O2 and C substrate diffusivity
Diffu<-function (gpo, wpo, taug, tauw){
  Dw_o2=1.4e-9*temp/298.0  #aqueous tracer diffusivity at 25
  Dg_o2=1.8e-5*(temp/273.0)**1.82 #oxygen diffusivity in gas phase
  henry_o2=3.2e-2*exp(-1500.*(1/temp-1/298.15))
  Dwo2= 0.5*(Dg_o2*taug*gpo*henry_o2+Dw_o2*tauw*wpo) #bulk aqueous molecular diffusivity as a colume weighted average between aquesous and gaseous phases
  Dw_s=6e-9 #oxygen diffusivity in water
  Dws=Dw_s*tauw*wpo#bulk substrate diffusivity (between the soil matrix and microsite)
  return(list(Dwo2=Dwo2, Dws=Dws))
}

gpo<-tortuo$gpo
wpo<-tortuo$wpo
taug<-tortuo$taug
tauw<-tortuo$tauw
Dw<-Diffu(gpo, wpo, taug, tauw)

#==calculating effective microbial substrate affinity for OC and O2 (Tang et al 2019)========
Kaff<-function(Ncell, gpo, wpo, taug, tauw, Dws,Dwg,mp){
  Bdens<-Ncell/Na #free microsite microbial abundance mol/m3
  Rm<-rc*(40*Ncell)^(1/3)  #microsite radius
  vm<-pi*4/3*Rm^3  #microsite volume 
  kw2<-k2_p*Npsite #(Npsite*rp+pi*rc) #cell specific uptake rate for OX, unit s-1
  Dw_s0 = 6e-9  #for claculating reference affinity
  Dw_g0 = 1.4e-9
  fin<-Npsite*rp/ (Npsite*rp+pi*rc)#interception probability (number of molecules that 1 mol cell will encounter and be able to intercept)
  ksw1<-4*pi*Dw_s0*rc*fin*Na   #substrate delivery parameter unit m3 mo-1 s-1
  kow1<-4*pi*Dw_g0*rc*fin*Na
  ksw0<-kw2/ksw1#reference affinity (Km used in MM kinetics in a well-mixed solution)
  kow0<-kw2/kow1#reference affinity (Km used in MM kinetics in a well-mixed solution)
  film<-exp(-13.65-0.857*log(mp/1000))    #calculting water film thickness
  ks_con<- (film/(Rm*Dw_s0*(Rm+film)) + 1/(Dws*(Rm+film)))*vm/(4*pi)#conductance coefficient
  ks_aff<-ksw0*(1+ks_con*ksw1*Ncell/Na/vm)
  ko_con<- (film/(Rm*Dw_g0*(Rm+film)) + 1/(Dwg*(Rm+film)))*vm/(4*pi)#conductance coefficient
  ko_aff<-kow0*(1+ko_con*kow1*Ncell/Na/vm)
  return(list(ksw0=ksw0, kow0=kow0,KaffOC=ks_aff, Kaffoxy=ko_aff))
}

mp<-rate$mp
Dws<-Dw$Dws
Dwg<-Dw$Dwo2
Ncell<-2.68e-10*rate$soc_ss*1e-6*6.02e23 
test<-Kaff(Ncell, gpo, wpo, taug, tauw, Dws,Dwg,mp)

rate$Keffs<-test$KaffOC*1e-3
rate$Keffo<-test$Kaffoxy*1e-3


kmdata<-data.frame(rs=rate$rs)
kmdata$kms_eff<-rate$Keffs
kmdata$kmo_eff<-rate$Keffo
ptest<-reshape2::melt(kmdata, id.vars=("rs"))

base_breaks <- function(n = 10){
    function(x) {
        axisTicks(log10(range(x, na.rm = TRUE)), log = TRUE, n = n)
    }
}

plot1<-ggplot(ptest, aes(x = rs,y=value, color=variable)) + 
  scale_x_continuous(name = "Relative saturation",limits = c(0,1)) +
  scale_y_continuous(name = "Effective substrate affinity",trans = log_trans(), breaks = base_breaks(),limits = c(1e-8,1e-5))+
  geom_point(aes(color=variable,linetype=variable), size=1)+
  scale_color_manual(values = c("#b2abd2","#2d004b"))+
  scale_linetype_manual(values=c("solid","solid"))

plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))
print(plot2)
pdf("effective_km.pdf", width=5.5, height=4)
plot2
dev.off() 
```

## new dataframe for simulations
This section put the calculated accessible soc/oxy and effective affinity parameter Keffs/Keffo back to the orginal data frame.

```{r newdata}
userate<-dplyr::select(rate, c(id, mp_fc,soc_ss, oxy_ss,Keffs, Keffo))
usesoil<-dplyr::select(fdata, c(id,rs,ra, mvol, mp,  bd, porosity, clay, org, mp_th))
newdata<-merge(usesoil, userate, by="id")
summary(newdata)
```
##Diffusion based moisture-respiration relationship
This section calculates mass transfer coefficients and analytical solution of diffusion-limited Michaelis-Menten microbial uptake kinetics
```{r diff-mmk}
#==========Diffusion based moisture-respiration relationship 
fh<-function(mvol,po,bd){
  Ds0=1.4e-9 #aqueous tracer diffusivity at 25 (m2/s)
  Dg0=2.1e-5 #oxygen diffusivity in gas
  fDg<-((po-mvol)/(po))^0.5 #gas phase relative diffusivity
  fDs<-((mvol)/(po))^0.5 #aqueous phase relative diffusivity
  H_o2<-1.3e-6  #mol/cm3/atm
  hs<-6/(mvol+bd*10)*Ds0*fDs/(0.00002^2) #DOM delivery (mass transfer rate in d-1)
  hg<-6/(mvol+bd*1)*Dg0*fDg*H_o2/(0.00002^2)  #DO delivery (mass transfer rate)
  return (list(fDs,fDg, hs, hg))}
mvol<-newdata$mvol
po<-newdata$porosity
bd<-newdata$bd
fhout<-fh(mvol, po, bd)
hs<-fhout[[3]]
hg<-fhout[[4]]

DiffMM<-function (hs,hg, soc, oxy, kmc, kmg,vmax){
  fm<-1-(log(newdata$mp/newdata$mp_fc)/log(newdata$mp_th/newdata$mp_fc))^1 #newdata$rs^(1/2.2) #microbial hydrological sensitivity
  ac<-kmc/(soc)
  bc<-vmax/hs/(soc)
  t1c<-(1-4*bc/(1+ac+bc)^2)^0.5
  F1c<-(1+ac+bc)/2/bc*(1-t1c)
  css<-F1c*kmc/(1-F1c)
  ag<-kmg/(oxy)
  bg<-vmax/hg/(oxy)
  t1g<-(1-4*bg/(1+ag+bg)^2)^0.5
  F1g<-(1+ag+bg)/2/bg*(1-t1g)
  oss<-F1g*kmg/(1-F1g)
  Ft<-fm*F1c*F1g*10^8
  rc<-ac/bc*ag/bg
  tc<-css/soc
  return(list(Ft,rc,tc))
}


soc<-newdata$soc_ss
oxy<-newdata$oxy_ss
kmo<-newdata$Keffo
kms<-newdata$Keffs

out1<-DiffMM(hs,hg, soc,oxy,1e-6,1e-6,1e-6) #Km_ref #1e-7,1e-7,2.5e-6, 1e-6,1e-6,5e-6
out2<-DiffMM(hs,hg, soc,oxy,kms,kmo,5e-6) #Km_eff
newdata$out1<-1.82*newdata$rs-1.14*newdata$rs^2#empirical (3.07RS-2.36RS^2) 1.98*rs-1.52*rs^2
newdata$out2<-1-(log(newdata$mp/newdata$mp_fc)/log(newdata$mp_th/newdata$mp_fc))^1#exp(-0.00021*newdata$mp)  #physiological  {1-(log(po/33)/log(15800/33))^2}
newdata$out3<-(1-(log(newdata$mp/newdata$mp_fc)/log(newdata$mp_th/newdata$mp_fc))^1) *fhout[[1]]*fhout[[2]]*soc*oxy #linear
newdata$out4<-out1[[1]]#km
newdata$out5<-out2[[1]]#variable km
summary(newdata)

```






```{r abtest}
ptest<-data.frame(rs=newdata$rs) 
ptest$rc<-out1[[2]]


#residual plot
plot1<-ggplot(ptest, aes(x = rs,y = rc)) +
  scale_x_continuous(name = expression(paste("Relative saturation")),limits = c(0,1)) +
  scale_y_continuous(name = 'a/b ratio')+geom_point(col="#642ba6", pch=1 , cex=1.6)+
  theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))
print(plot1)
pdf("incu_abratio.pdf", width=5.5, height=4)
plot2
dev.off() 

```

## Rescale simulations
Rescale simulation results to 0-1 scale (because vmaxs are not optimized to ra measurements)

```{r rescale}

mpdata<-newdata
colnames(mpdata)[16]<-("pred_new")
mpdata$scaled_new<-0

#check unique id======
unique(mpdata$id)

id_1<-subset(mpdata, id=="Bare")
id_1$scaled_new<-id_1$pred_new/max(id_1$pred_new)

id_2<-subset(mpdata, id=="Cedao")
id_2$scaled_new<-id_2$pred_new/max(id_2$pred_new)

id_3<-subset(mpdata, id=="Cedo")
id_3$scaled_new<-id_3$pred_new/max(id_3$pred_new)

id_4<-subset(mpdata, id=="droughtCK")
id_4$scaled_new<-id_4$pred_new/max(id_4$pred_new)

id_5<-subset(mpdata, id=="droughtPE3")
id_5$scaled_new<-id_5$pred_new/max(id_5$pred_new)

id_6<-subset(mpdata, id=="droughtPE5")
id_6$scaled_new<-id_6$pred_new/max(id_6$pred_new)

id_7<-subset(mpdata, id=="droughtPE7")
id_7$scaled_new<-id_7$pred_new/max(id_7$pred_new)

id_8<-subset(mpdata, id=="Fog1")
id_8$scaled_new<-id_8$pred_new/max(id_8$pred_new)

id_9<-subset(mpdata, id=="Fog2")
id_9$scaled_new<-id_9$pred_new/max(id_9$pred_new)

id_10<-subset(mpdata, id=="GE")
id_10$scaled_new<-id_10$pred_new/max(id_10$pred_new)

id_11<-subset(mpdata, id=="GiganteAd")
id_11$scaled_new<-id_11$pred_new/max(id_11$pred_new)

id_12<-subset(mpdata, id=="GiganteCT")
id_12$scaled_new<-id_12$pred_new/max(id_12$pred_new)

id_13<-subset(mpdata, id=="GiganteRe")
id_13$scaled_new<-id_13$pred_new/max(id_13$pred_new)

id_14<-subset(mpdata, id=="HC")
id_14$scaled_new<-id_14$pred_new/max(id_14$pred_new)

id_15<-subset(mpdata, id=="LC")
id_15$scaled_new<-id_15$pred_new/max(id_15$pred_new)

id_16<-subset(mpdata, id=="MC")
id_16$scaled_new<-id_16$pred_new/max(id_16$pred_new)

id_17<-subset(mpdata, id=="mix")
id_17$scaled_new<-id_17$pred_new/max(id_17$pred_new)

id_18<-subset(mpdata, id=="OakOA")
id_18$scaled_new<-id_18$pred_new/max(id_18$pred_new)

id_19<-subset(mpdata, id=="OakUT")
id_19$scaled_new<-id_19$pred_new/max(id_19$pred_new)

id_20<-subset(mpdata, id=="open")
id_20$scaled_new<-id_20$pred_new/max(id_20$pred_new)

id_21<-subset(mpdata, id=="Pineforest")
id_21$scaled_new<-id_21$pred_new/max(id_21$pred_new)

id_22<-subset(mpdata, id=="Rain1")
id_22$scaled_new<-id_22$pred_new/max(id_22$pred_new)

id_23<-subset(mpdata, id=="Rain2")
id_23$scaled_new<-id_23$pred_new/max(id_23$pred_new)

id_24<-subset(mpdata, id=="Range")
id_24$scaled_new<-id_24$pred_new/max(id_24$pred_new)

id_25<-subset(mpdata, id=="RR")
id_25$scaled_new<-id_25$pred_new/max(id_25$pred_new)

id_26<-subset(mpdata, id=="RS")
id_26$scaled_new<-id_26$pred_new/max(id_26$pred_new)

id_27<-subset(mpdata, id=="ST")
id_27$scaled_new<-id_27$pred_new/max(id_27$pred_new)

id_28<-subset(mpdata, id=="Teak")
id_28$scaled_new<-id_28$pred_new/max(id_28$pred_new)

id_29<-subset(mpdata, id=="TibetCK")
id_29$scaled_new<-id_29$pred_new/max(id_29$pred_new)

id_30<-subset(mpdata, id=="TibetW1")
id_30$scaled_new<-id_30$pred_new/max(id_30$pred_new)

id_31<-subset(mpdata, id=="TibetW2")
id_31$scaled_new<-id_31$pred_new/max(id_31$pred_new)

id_32<-subset(mpdata, id=="under")
id_32$scaled_new<-id_32$pred_new/max(id_32$pred_new)

id_33<-subset(mpdata, id=="WythamAd")
id_33$scaled_new<-id_33$pred_new/max(id_33$pred_new)

id_34<-subset(mpdata, id=="WythamCT")
id_34$scaled_new<-id_34$pred_new/max(id_34$pred_new)

id_35<-subset(mpdata, id=="WythamRe")
id_35$scaled_new<-id_35$pred_new/max(id_35$pred_new)

id_36<-subset(mpdata, id=="YL")
id_36$scaled_new<-id_36$pred_new/max(id_36$pred_new)


new_scaled<-rbind(id_1,id_2,id_3,id_4,id_5,id_6,id_7,id_8,id_9,id_10,id_11,id_12,id_13,id_14,id_15,id_16,id_17,id_18,id_19,id_20,
                  id_21,id_22,id_23,id_24,id_25,id_26,id_27,id_28,id_29,id_30,id_31,id_32,id_33,id_34,id_35,id_36)
#=====
newdata$out1_s<-new_scaled$scaled_new

mpdata<-newdata
colnames(mpdata)[17]<-("pred_new")
mpdata$scaled_new<-0

#check unique id======
unique(mpdata$id)

id_1<-subset(mpdata, id=="Bare")
id_1$scaled_new<-id_1$pred_new/max(id_1$pred_new)

id_2<-subset(mpdata, id=="Cedao")
id_2$scaled_new<-id_2$pred_new/max(id_2$pred_new)

id_3<-subset(mpdata, id=="Cedo")
id_3$scaled_new<-id_3$pred_new/max(id_3$pred_new)

id_4<-subset(mpdata, id=="droughtCK")
id_4$scaled_new<-id_4$pred_new/max(id_4$pred_new)

id_5<-subset(mpdata, id=="droughtPE3")
id_5$scaled_new<-id_5$pred_new/max(id_5$pred_new)

id_6<-subset(mpdata, id=="droughtPE5")
id_6$scaled_new<-id_6$pred_new/max(id_6$pred_new)

id_7<-subset(mpdata, id=="droughtPE7")
id_7$scaled_new<-id_7$pred_new/max(id_7$pred_new)

id_8<-subset(mpdata, id=="Fog1")
id_8$scaled_new<-id_8$pred_new/max(id_8$pred_new)

id_9<-subset(mpdata, id=="Fog2")
id_9$scaled_new<-id_9$pred_new/max(id_9$pred_new)

id_10<-subset(mpdata, id=="GE")
id_10$scaled_new<-id_10$pred_new/max(id_10$pred_new)

id_11<-subset(mpdata, id=="GiganteAd")
id_11$scaled_new<-id_11$pred_new/max(id_11$pred_new)

id_12<-subset(mpdata, id=="GiganteCT")
id_12$scaled_new<-id_12$pred_new/max(id_12$pred_new)

id_13<-subset(mpdata, id=="GiganteRe")
id_13$scaled_new<-id_13$pred_new/max(id_13$pred_new)

id_14<-subset(mpdata, id=="HC")
id_14$scaled_new<-id_14$pred_new/max(id_14$pred_new)

id_15<-subset(mpdata, id=="LC")
id_15$scaled_new<-id_15$pred_new/max(id_15$pred_new)

id_16<-subset(mpdata, id=="MC")
id_16$scaled_new<-id_16$pred_new/max(id_16$pred_new)

id_17<-subset(mpdata, id=="mix")
id_17$scaled_new<-id_17$pred_new/max(id_17$pred_new)

id_18<-subset(mpdata, id=="OakOA")
id_18$scaled_new<-id_18$pred_new/max(id_18$pred_new)

id_19<-subset(mpdata, id=="OakUT")
id_19$scaled_new<-id_19$pred_new/max(id_19$pred_new)

id_20<-subset(mpdata, id=="open")
id_20$scaled_new<-id_20$pred_new/max(id_20$pred_new)

id_21<-subset(mpdata, id=="Pineforest")
id_21$scaled_new<-id_21$pred_new/max(id_21$pred_new)

id_22<-subset(mpdata, id=="Rain1")
id_22$scaled_new<-id_22$pred_new/max(id_22$pred_new)

id_23<-subset(mpdata, id=="Rain2")
id_23$scaled_new<-id_23$pred_new/max(id_23$pred_new)

id_24<-subset(mpdata, id=="Range")
id_24$scaled_new<-id_24$pred_new/max(id_24$pred_new)

id_25<-subset(mpdata, id=="RR")
id_25$scaled_new<-id_25$pred_new/max(id_25$pred_new)

id_26<-subset(mpdata, id=="RS")
id_26$scaled_new<-id_26$pred_new/max(id_26$pred_new)

id_27<-subset(mpdata, id=="ST")
id_27$scaled_new<-id_27$pred_new/max(id_27$pred_new)

id_28<-subset(mpdata, id=="Teak")
id_28$scaled_new<-id_28$pred_new/max(id_28$pred_new)

id_29<-subset(mpdata, id=="TibetCK")
id_29$scaled_new<-id_29$pred_new/max(id_29$pred_new)

id_30<-subset(mpdata, id=="TibetW1")
id_30$scaled_new<-id_30$pred_new/max(id_30$pred_new)

id_31<-subset(mpdata, id=="TibetW2")
id_31$scaled_new<-id_31$pred_new/max(id_31$pred_new)

id_32<-subset(mpdata, id=="under")
id_32$scaled_new<-id_32$pred_new/max(id_32$pred_new)

id_33<-subset(mpdata, id=="WythamAd")
id_33$scaled_new<-id_33$pred_new/max(id_33$pred_new)

id_34<-subset(mpdata, id=="WythamCT")
id_34$scaled_new<-id_34$pred_new/max(id_34$pred_new)

id_35<-subset(mpdata, id=="WythamRe")
id_35$scaled_new<-id_35$pred_new/max(id_35$pred_new)

id_36<-subset(mpdata, id=="YL")
id_36$scaled_new<-id_36$pred_new/max(id_36$pred_new)


new_scaled<-rbind(id_1,id_2,id_3,id_4,id_5,id_6,id_7,id_8,id_9,id_10,id_11,id_12,id_13,id_14,id_15,id_16,id_17,id_18,id_19,id_20,
                  id_21,id_22,id_23,id_24,id_25,id_26,id_27,id_28,id_29,id_30,id_31,id_32,id_33,id_34,id_35,id_36)
#=====
newdata$out2_s<-new_scaled$scaled_new

mpdata<-newdata
colnames(mpdata)[18]<-("pred_new")
mpdata$scaled_new<-0

#check unique id======
unique(mpdata$id)

id_1<-subset(mpdata, id=="Bare")
id_1$scaled_new<-id_1$pred_new/max(id_1$pred_new)

id_2<-subset(mpdata, id=="Cedao")
id_2$scaled_new<-id_2$pred_new/max(id_2$pred_new)

id_3<-subset(mpdata, id=="Cedo")
id_3$scaled_new<-id_3$pred_new/max(id_3$pred_new)

id_4<-subset(mpdata, id=="droughtCK")
id_4$scaled_new<-id_4$pred_new/max(id_4$pred_new)

id_5<-subset(mpdata, id=="droughtPE3")
id_5$scaled_new<-id_5$pred_new/max(id_5$pred_new)

id_6<-subset(mpdata, id=="droughtPE5")
id_6$scaled_new<-id_6$pred_new/max(id_6$pred_new)

id_7<-subset(mpdata, id=="droughtPE7")
id_7$scaled_new<-id_7$pred_new/max(id_7$pred_new)

id_8<-subset(mpdata, id=="Fog1")
id_8$scaled_new<-id_8$pred_new/max(id_8$pred_new)

id_9<-subset(mpdata, id=="Fog2")
id_9$scaled_new<-id_9$pred_new/max(id_9$pred_new)

id_10<-subset(mpdata, id=="GE")
id_10$scaled_new<-id_10$pred_new/max(id_10$pred_new)

id_11<-subset(mpdata, id=="GiganteAd")
id_11$scaled_new<-id_11$pred_new/max(id_11$pred_new)

id_12<-subset(mpdata, id=="GiganteCT")
id_12$scaled_new<-id_12$pred_new/max(id_12$pred_new)

id_13<-subset(mpdata, id=="GiganteRe")
id_13$scaled_new<-id_13$pred_new/max(id_13$pred_new)

id_14<-subset(mpdata, id=="HC")
id_14$scaled_new<-id_14$pred_new/max(id_14$pred_new)

id_15<-subset(mpdata, id=="LC")
id_15$scaled_new<-id_15$pred_new/max(id_15$pred_new)

id_16<-subset(mpdata, id=="MC")
id_16$scaled_new<-id_16$pred_new/max(id_16$pred_new)

id_17<-subset(mpdata, id=="mix")
id_17$scaled_new<-id_17$pred_new/max(id_17$pred_new)

id_18<-subset(mpdata, id=="OakOA")
id_18$scaled_new<-id_18$pred_new/max(id_18$pred_new)

id_19<-subset(mpdata, id=="OakUT")
id_19$scaled_new<-id_19$pred_new/max(id_19$pred_new)

id_20<-subset(mpdata, id=="open")
id_20$scaled_new<-id_20$pred_new/max(id_20$pred_new)

id_21<-subset(mpdata, id=="Pineforest")
id_21$scaled_new<-id_21$pred_new/max(id_21$pred_new)

id_22<-subset(mpdata, id=="Rain1")
id_22$scaled_new<-id_22$pred_new/max(id_22$pred_new)

id_23<-subset(mpdata, id=="Rain2")
id_23$scaled_new<-id_23$pred_new/max(id_23$pred_new)

id_24<-subset(mpdata, id=="Range")
id_24$scaled_new<-id_24$pred_new/max(id_24$pred_new)

id_25<-subset(mpdata, id=="RR")
id_25$scaled_new<-id_25$pred_new/max(id_25$pred_new)

id_26<-subset(mpdata, id=="RS")
id_26$scaled_new<-id_26$pred_new/max(id_26$pred_new)

id_27<-subset(mpdata, id=="ST")
id_27$scaled_new<-id_27$pred_new/max(id_27$pred_new)

id_28<-subset(mpdata, id=="Teak")
id_28$scaled_new<-id_28$pred_new/max(id_28$pred_new)

id_29<-subset(mpdata, id=="TibetCK")
id_29$scaled_new<-id_29$pred_new/max(id_29$pred_new)

id_30<-subset(mpdata, id=="TibetW1")
id_30$scaled_new<-id_30$pred_new/max(id_30$pred_new)

id_31<-subset(mpdata, id=="TibetW2")
id_31$scaled_new<-id_31$pred_new/max(id_31$pred_new)

id_32<-subset(mpdata, id=="under")
id_32$scaled_new<-id_32$pred_new/max(id_32$pred_new)

id_33<-subset(mpdata, id=="WythamAd")
id_33$scaled_new<-id_33$pred_new/max(id_33$pred_new)

id_34<-subset(mpdata, id=="WythamCT")
id_34$scaled_new<-id_34$pred_new/max(id_34$pred_new)

id_35<-subset(mpdata, id=="WythamRe")
id_35$scaled_new<-id_35$pred_new/max(id_35$pred_new)

id_36<-subset(mpdata, id=="YL")
id_36$scaled_new<-id_36$pred_new/max(id_36$pred_new)

new_scaled<-rbind(id_1,id_2,id_3,id_4,id_5,id_6,id_7,id_8,id_9,id_10,id_11,id_12,id_13,id_14,id_15,id_16,id_17,id_18,id_19,id_20,
                  id_21,id_22,id_23,id_24,id_25,id_26,id_27,id_28,id_29,id_30,id_31,id_32,id_33,id_34,id_35,id_36)
#=====
newdata$out3_s<-new_scaled$scaled_new



mpdata<-newdata
colnames(mpdata)[19]<-("pred_new")
mpdata$scaled_new<-0

#check unique id======
unique(mpdata$id)

id_1<-subset(mpdata, id=="Bare")
id_1$scaled_new<-id_1$pred_new/max(id_1$pred_new)

id_2<-subset(mpdata, id=="Cedao")
id_2$scaled_new<-id_2$pred_new/max(id_2$pred_new)

id_3<-subset(mpdata, id=="Cedo")
id_3$scaled_new<-id_3$pred_new/max(id_3$pred_new)

id_4<-subset(mpdata, id=="droughtCK")
id_4$scaled_new<-id_4$pred_new/max(id_4$pred_new)

id_5<-subset(mpdata, id=="droughtPE3")
id_5$scaled_new<-id_5$pred_new/max(id_5$pred_new)

id_6<-subset(mpdata, id=="droughtPE5")
id_6$scaled_new<-id_6$pred_new/max(id_6$pred_new)

id_7<-subset(mpdata, id=="droughtPE7")
id_7$scaled_new<-id_7$pred_new/max(id_7$pred_new)

id_8<-subset(mpdata, id=="Fog1")
id_8$scaled_new<-id_8$pred_new/max(id_8$pred_new)

id_9<-subset(mpdata, id=="Fog2")
id_9$scaled_new<-id_9$pred_new/max(id_9$pred_new)

id_10<-subset(mpdata, id=="GE")
id_10$scaled_new<-id_10$pred_new/max(id_10$pred_new)

id_11<-subset(mpdata, id=="GiganteAd")
id_11$scaled_new<-id_11$pred_new/max(id_11$pred_new)

id_12<-subset(mpdata, id=="GiganteCT")
id_12$scaled_new<-id_12$pred_new/max(id_12$pred_new)

id_13<-subset(mpdata, id=="GiganteRe")
id_13$scaled_new<-id_13$pred_new/max(id_13$pred_new)

id_14<-subset(mpdata, id=="HC")
id_14$scaled_new<-id_14$pred_new/max(id_14$pred_new)

id_15<-subset(mpdata, id=="LC")
id_15$scaled_new<-id_15$pred_new/max(id_15$pred_new)

id_16<-subset(mpdata, id=="MC")
id_16$scaled_new<-id_16$pred_new/max(id_16$pred_new)

id_17<-subset(mpdata, id=="mix")
id_17$scaled_new<-id_17$pred_new/max(id_17$pred_new)

id_18<-subset(mpdata, id=="OakOA")
id_18$scaled_new<-id_18$pred_new/max(id_18$pred_new)

id_19<-subset(mpdata, id=="OakUT")
id_19$scaled_new<-id_19$pred_new/max(id_19$pred_new)

id_20<-subset(mpdata, id=="open")
id_20$scaled_new<-id_20$pred_new/max(id_20$pred_new)

id_21<-subset(mpdata, id=="Pineforest")
id_21$scaled_new<-id_21$pred_new/max(id_21$pred_new)

id_22<-subset(mpdata, id=="Rain1")
id_22$scaled_new<-id_22$pred_new/max(id_22$pred_new)

id_23<-subset(mpdata, id=="Rain2")
id_23$scaled_new<-id_23$pred_new/max(id_23$pred_new)

id_24<-subset(mpdata, id=="Range")
id_24$scaled_new<-id_24$pred_new/max(id_24$pred_new)

id_25<-subset(mpdata, id=="RR")
id_25$scaled_new<-id_25$pred_new/max(id_25$pred_new)

id_26<-subset(mpdata, id=="RS")
id_26$scaled_new<-id_26$pred_new/max(id_26$pred_new)

id_27<-subset(mpdata, id=="ST")
id_27$scaled_new<-id_27$pred_new/max(id_27$pred_new)

id_28<-subset(mpdata, id=="Teak")
id_28$scaled_new<-id_28$pred_new/max(id_28$pred_new)

id_29<-subset(mpdata, id=="TibetCK")
id_29$scaled_new<-id_29$pred_new/max(id_29$pred_new)

id_30<-subset(mpdata, id=="TibetW1")
id_30$scaled_new<-id_30$pred_new/max(id_30$pred_new)

id_31<-subset(mpdata, id=="TibetW2")
id_31$scaled_new<-id_31$pred_new/max(id_31$pred_new)

id_32<-subset(mpdata, id=="under")
id_32$scaled_new<-id_32$pred_new/max(id_32$pred_new)

id_33<-subset(mpdata, id=="WythamAd")
id_33$scaled_new<-id_33$pred_new/max(id_33$pred_new)

id_34<-subset(mpdata, id=="WythamCT")
id_34$scaled_new<-id_34$pred_new/max(id_34$pred_new)

id_35<-subset(mpdata, id=="WythamRe")
id_35$scaled_new<-id_35$pred_new/max(id_35$pred_new)

id_36<-subset(mpdata, id=="YL")
id_36$scaled_new<-id_36$pred_new/max(id_36$pred_new)

new_scaled<-rbind(id_1,id_2,id_3,id_4,id_5,id_6,id_7,id_8,id_9,id_10,id_11,id_12,id_13,id_14,id_15,id_16,id_17,id_18,id_19,id_20,
                  id_21,id_22,id_23,id_24,id_25,id_26,id_27,id_28,id_29,id_30,id_31,id_32,id_33,id_34,id_35,id_36)
#=====
newdata$out4_s<-new_scaled$scaled_new


mpdata<-newdata
colnames(mpdata)[20]<-("pred_new")
mpdata$scaled_new<-0

#check unique id======
unique(mpdata$id)

id_1<-subset(mpdata, id=="Bare")
id_1$scaled_new<-id_1$pred_new/max(id_1$pred_new)

id_2<-subset(mpdata, id=="Cedao")
id_2$scaled_new<-id_2$pred_new/max(id_2$pred_new)

id_3<-subset(mpdata, id=="Cedo")
id_3$scaled_new<-id_3$pred_new/max(id_3$pred_new)

id_4<-subset(mpdata, id=="droughtCK")
id_4$scaled_new<-id_4$pred_new/max(id_4$pred_new)

id_5<-subset(mpdata, id=="droughtPE3")
id_5$scaled_new<-id_5$pred_new/max(id_5$pred_new)

id_6<-subset(mpdata, id=="droughtPE5")
id_6$scaled_new<-id_6$pred_new/max(id_6$pred_new)

id_7<-subset(mpdata, id=="droughtPE7")
id_7$scaled_new<-id_7$pred_new/max(id_7$pred_new)

id_8<-subset(mpdata, id=="Fog1")
id_8$scaled_new<-id_8$pred_new/max(id_8$pred_new)

id_9<-subset(mpdata, id=="Fog2")
id_9$scaled_new<-id_9$pred_new/max(id_9$pred_new)

id_10<-subset(mpdata, id=="GE")
id_10$scaled_new<-id_10$pred_new/max(id_10$pred_new)

id_11<-subset(mpdata, id=="GiganteAd")
id_11$scaled_new<-id_11$pred_new/max(id_11$pred_new)

id_12<-subset(mpdata, id=="GiganteCT")
id_12$scaled_new<-id_12$pred_new/max(id_12$pred_new)

id_13<-subset(mpdata, id=="GiganteRe")
id_13$scaled_new<-id_13$pred_new/max(id_13$pred_new)

id_14<-subset(mpdata, id=="HC")
id_14$scaled_new<-id_14$pred_new/max(id_14$pred_new)

id_15<-subset(mpdata, id=="LC")
id_15$scaled_new<-id_15$pred_new/max(id_15$pred_new)

id_16<-subset(mpdata, id=="MC")
id_16$scaled_new<-id_16$pred_new/max(id_16$pred_new)

id_17<-subset(mpdata, id=="mix")
id_17$scaled_new<-id_17$pred_new/max(id_17$pred_new)

id_18<-subset(mpdata, id=="OakOA")
id_18$scaled_new<-id_18$pred_new/max(id_18$pred_new)

id_19<-subset(mpdata, id=="OakUT")
id_19$scaled_new<-id_19$pred_new/max(id_19$pred_new)

id_20<-subset(mpdata, id=="open")
id_20$scaled_new<-id_20$pred_new/max(id_20$pred_new)

id_21<-subset(mpdata, id=="Pineforest")
id_21$scaled_new<-id_21$pred_new/max(id_21$pred_new)

id_22<-subset(mpdata, id=="Rain1")
id_22$scaled_new<-id_22$pred_new/max(id_22$pred_new)

id_23<-subset(mpdata, id=="Rain2")
id_23$scaled_new<-id_23$pred_new/max(id_23$pred_new)

id_24<-subset(mpdata, id=="Range")
id_24$scaled_new<-id_24$pred_new/max(id_24$pred_new)

id_25<-subset(mpdata, id=="RR")
id_25$scaled_new<-id_25$pred_new/max(id_25$pred_new)

id_26<-subset(mpdata, id=="RS")
id_26$scaled_new<-id_26$pred_new/max(id_26$pred_new)

id_27<-subset(mpdata, id=="ST")
id_27$scaled_new<-id_27$pred_new/max(id_27$pred_new)

id_28<-subset(mpdata, id=="Teak")
id_28$scaled_new<-id_28$pred_new/max(id_28$pred_new)

id_29<-subset(mpdata, id=="TibetCK")
id_29$scaled_new<-id_29$pred_new/max(id_29$pred_new)

id_30<-subset(mpdata, id=="TibetW1")
id_30$scaled_new<-id_30$pred_new/max(id_30$pred_new)

id_31<-subset(mpdata, id=="TibetW2")
id_31$scaled_new<-id_31$pred_new/max(id_31$pred_new)

id_32<-subset(mpdata, id=="under")
id_32$scaled_new<-id_32$pred_new/max(id_32$pred_new)

id_33<-subset(mpdata, id=="WythamAd")
id_33$scaled_new<-id_33$pred_new/max(id_33$pred_new)

id_34<-subset(mpdata, id=="WythamCT")
id_34$scaled_new<-id_34$pred_new/max(id_34$pred_new)

id_35<-subset(mpdata, id=="WythamRe")
id_35$scaled_new<-id_35$pred_new/max(id_35$pred_new)

id_36<-subset(mpdata, id=="YL")
id_36$scaled_new<-id_36$pred_new/max(id_36$pred_new)

new_scaled<-rbind(id_1,id_2,id_3,id_4,id_5,id_6,id_7,id_8,id_9,id_10,id_11,id_12,id_13,id_14,id_15,id_16,id_17,id_18,id_19,id_20,
                  id_21,id_22,id_23,id_24,id_25,id_26,id_27,id_28,id_29,id_30,id_31,id_32,id_33,id_34,id_35,id_36)
#=====
newdata$out5_s<-new_scaled$scaled_new

```

```{r plot}
testplot<-data.frame(RS=newdata$rs)
testplot$Obv<-newdata$ra
testplot$Emp<-newdata$out1
#testplot$Physiology<-newdata$out2_s
testplot$Lin<-newdata$out3_s
testplot$MMK<-newdata$out4_s
testplot$MMK_eff<-newdata$out5_s
#create matrix plot
panel.hist <- function(x, ...) {
  usr <- par("usr"); on.exit(par(usr))
  par(usr = c(usr[1:2], 0, 1) )
  h <- hist(x, breaks=20,plot = FALSE)
  breaks <- h$breaks; nB <- length(breaks)
  y <- h$counts; y <- y/max(y)
  rect(breaks[-nB], 0, breaks[-1], y, ...)
}

#reg <- function(x, y, col) abline(lm(y~x), col=col) 

panel.lm =  function (x, y, col = par("col"), bg = NA, pch = par("pch"), 
    cex = 1, col.smooth = "black", span = 2/3, iter = 3, ...)  {
    points(x, y, pch = pch, col = col, bg = bg, cex = cex)
    ok <- is.finite(x) & is.finite(y)
    if (any(ok)) reg(x[ok], y[ok], col.smooth)
}

pairs(testplot, pch = 19,  cex = 0.5, col=c("#b35806"),diag.panel=NULL,
      upper.panel=NULL)

dev.print(pdf,        # copies the plot to a the PDF file
          "Corr_matrix3.pdf",width=7, height=5.8)     

```
```{r gradient}

testplot<-data.frame(rs=newdata$rs)
testplot$id<-newdata$id
testplot$Obv<-newdata$ra
testplot$Emp<-newdata$out1
#testplot$Physiology<-newdata$out2_s
testplot$Lin<-newdata$out3_s
testplot$MMK<-newdata$out4_s
testplot$MMK_eff<-newdata$out5_s
testplot$clay<-newdata$clay


#plot observations along clay and soc gradient
plot1<-ggplot(testplot, aes(x = rs,y = MMK_eff, group = id)) +
  scale_x_continuous(name = expression(paste("Relative saturation")),limits = c(0,1)) +
  scale_y_continuous(name = 'Relative activity ',limits = c(0,1)) +
  geom_line(aes(color=clay),size=1.6, shape=1,stroke=1)+theme_bw()+
  scale_color_gradientn(colors=c("#b35806", "#fee0b6", "#542788"), values=c(0,1))+theme(text = element_text(size=16)) 
plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))#co
plot2
#pdf("field_mmk_clay.pdf", width=5.5, height=4)
#plot2
#dev.off() 
```

##result statistics and visualization
```{r visual}

ptest<-data.frame(rs=newdata$rs) 
ptest$res<-newdata$out4_s-newdata$ra

ave<-mean(newdata$ra)
ssres<-sum(ptest$res^2)
sstot<-sum((newdata$ra-ave)^2)  
RR<-1-ssres/sstot
RSME<-(ssres^2/length(newdata$ra))^1/2
paste0("R-square = ", RR) #print R-square value

paste0("RSME = ", RSME)


simple.fit = lm(res~rs, data=ptest)
summary(simple.fit)


#residual plot
plot1<-ggplot(ptest, aes(x = rs,y = res)) +
  scale_x_continuous(name = expression(paste("Relative saturation")),limits = c(0,1)) +
  scale_y_continuous(name = 'Residual',limits = c(-1,1))+geom_point(col="#642ba6", pch=1 , cex=1.6)+geom_hline(yintercept=0, color="black", linetype="dashed",size=0.5)+
  geom_smooth(method=lm, color="tomato1", size=1, fill="tomato1", se=TRUE)

plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))
print(plot2)

```


```{r line}


ptest<-data.frame(ra=newdata$ra) 


#ptest$lin<-newdata$out3_s
ptest$mmk<-newdata$out4_s
ptest$emp<-newdata$out1/0.73
pdiff<-reshape2::melt(ptest, id.vars=("ra"))
summary(pdiff)
x<-ptest$ra
y<-ptest$mmk
fit1<-lm(y~x)
summary(fit1)
  
plot1<-ggplot(pdiff, aes(x=ra, y=value))+geom_point(aes(color=variable),cex=1.2,shape=1, alpha=0.9)+
    geom_smooth(aes(color=variable, fill=variable),method = lm, se = TRUE, size=1)+
    geom_abline(slope=1,intercept=0, color="black",linetype="dashed",size=0.6)+
    scale_color_manual(values = c("#b35806","#8073ac","#7f3b08"))+
    scale_fill_manual(values = c("#b35806","#8073ac","#7f3b08"))+
    scale_x_continuous(name = 'Observation',limits = c(0,1))+
    scale_y_continuous(name = 'Prediction',limits = c(0,1))+
    theme(text = element_text(size=16)) 
plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))
print(plot2)

#pdf("field_pred_size1.pdf", width=5, height=3.5)
#plot2
#dev.off() 

```


#residual plot
```{r line}


ptest<-data.frame(ra=newdata$ra) 

ptest$mmk<-newdata$out4_s
#ptest$lin<-newdata$out3_s
ptest$lin<-newdata$out3_s
pdiff<-reshape2::melt(ptest, id.vars=("ra"))
summary(pdiff)
x<-ptest$ra
y<-ptest$lin
fit1<-lm(y~x)
summary(fit1)
  
plot1<-ggplot(pdiff, aes(x=ra, y=value))+geom_point(aes(color=variable),cex=1,shape=1, alpha=0.9)+
    geom_smooth(aes(color=variable, fill=variable),method = lm, se = TRUE, size=1)+
    geom_abline(slope=1,intercept=0, color="black",linetype="dashed",size=0.6)+
    scale_color_manual(values = c("#b35806","#fdb863","#7f3b08"))+
    scale_fill_manual(values =c("#b35806","#fdb863","#7f3b08"))+
    scale_x_continuous(name = 'Observation',limits = c(0,1))+
    scale_y_continuous(name = 'Prediction',limits = c(0,1))+
    theme(text = element_text(size=16)) 
plot2<-plot1+theme_linedraw()+theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+theme(text = element_text(size=16))
print(plot2)

pdf("field_pred_new2.pdf", width=5.5, height=4)
plot2
dev.off() 

```
